---
title: "Exam 2"
format: html
---

### 2. Background

## 2.2 Load Packages
```{r, message = FALSE}
library(tidytuesdayR)
library(tidyverse)
library(rnaturalearth)
library(sf)
```

## 2.3 Load Data
```{r}
tuesdata <- tt_load('2020-02-18')
fc <- tuesdata$food_consumption
```
## 2.4 Inspect Data
```{r}
str(fc)
```
```{r}
head(fc)
```

## 2.5 Inspect country Variable 

## 2.6 Inspect food_category Variable

```{r}
fc |>
  distinct(food_category)
```
The problem with these values is that some of the categories overlap, like eggs and poultry and that the names are long.

## 2.7 Fix food_category Variable
```{r}
fcc <- fc|>
  mutate(food_category = str_replace(food_category, "Lamb & Goat", "Lamb"))|>
  mutate(food_category = str_replace(food_category, "Milk - inc. cheese", "Dairy")) |>
  mutate(food_category = str_replace(food_category, "Wheat and Wheat Products", "Wheat")) |>
  mutate(food_category = str_replace(food_category, "Nuts inc. Peanut Butter", "Nuts"))
```


## 2.8 Re-inspect food_category Variable

```{r}
fcc |>
  distinct(food_category)
```

### 3. Most Consuming Countries
# Research Question 1

Which 5 countries consume the most food? 

## 3.1 Expected Shape 

## 3.2 Steps 

1. Select Country and Consumption 
2. Group by Country 
3. Arrange by Consumption
4. Filter top 5

```{r}
fcc |>
  select(country, consumption) |>
  group_by(country)|>
  arrange(desc(consumption)) |>
  filter(consumption > 303.71 )
```


## 3.3 Viz

```{r}
fcc |>
  select(country, consumption) |>
  group_by(country)|>
  arrange(desc(consumption)) |>
  filter(consumption > 303.71 )|>
  ggplot(aes(x = country, y = consumption)) +
  geom_point()
```

### Most Consuming Countries of Each Food
# Research Question 2

Which top 5 countries consume each food are? 

## 4.1 Expected Shape 
## 4.2 Steps

1. Select Country, Consumption, and Food Category
2. Group by Country and Food Category
```{r}
fcc |>
  select(food_category,country, consumption)|>
  group_by(food_category, country)
```


## 4.3 Viz
```{r}
#fcc |>
 # select(food_category,country, consumption)|>
 # group_by(food_category, country) |>
 # ggplot(aes(x = consumption, fill = country)) +
  #geom_density()
```

