---
title: "Exam 1"
format: html
---

### Background

## 1.2 Grand Research Question 

**What does the consumption of each food category in each country look like?**

### 2. Install Packages

### 3. Load Packages
```{r}
library(tidytuesdayR)
library(tidyverse)
```

There are 9 packages loaded when loading the *tidyverse* package

### 4. Get Data
```{r}
tuesdata <- tt_load('2020-02-18')
fc <- tuesdata$food_consumption
```

### 5. Understand Data
```{r}
# Look at first 6 rows
head(fc)

# Find dimension of data set
dim(fc)

# Look at names of columns 
names(fc)

```
### 6. Explore Data
## 6.3 Observations

- What are the units of observations? 
Country and food_catefory

- How many food categories are there?
11

- How many countries are there?
130

### 7. Understand Variables Individually 

- How many variables does the grand research question involve?
3

### 8. Understand Consumption 

List appropriate plot for each bivariate viz and what should go into their aesthetic details

- Overall Food Consumption/ Food Category
Density Plot + Faceted

- Overall Food Consumption/ Country
Density Plot + Faceted

```{r}
#| fig-height: 22
#| fig-width: 11
ggplot(fc, aes(x = consumption, fill = food_category)) +
  geom_density()
```

### 9. Answering the Grand RQ

- Which of these plots is the most appropriate one? Why?
The map will be most clear for showing dif. food categories consumption by country

### 10. Beyond Viz
## 10.1 Effectiveness
List a minimum of **five** concepts that you should apply to your final viz to make it more effective?

1. Axis labels
2. Caption
3. Alt-text
4. Color-blind friendly color palette
5. Title
6. Legend title

## 10.2 Additional Questions
List **two** additional questions, new or follow-up that you would like to answer based on this dataset?

1. How consumption of dif. food categories correlates to CO2 emmission?
2. How does geographic location affect consumption of food types? 

### Finalize Work
