---
title: "Factors"
number-sections: true
execute: 
  warning: false
fig-height: 2.75
fig-width: 4.25
fig-env: 'figure'
fig-pos: 'h'
fig-align: center
code-fold: false
---


### 12.3 Exercises

```{r}
library(tidyverse)

# Loading Datasets! 
grades <- read.csv("https://mac-stat.github.io/data/grades.csv") |> 
  distinct(sid, sessionID, .keep_all = TRUE)

head(grades)

courses <- read.csv("https://mac-stat.github.io/data/courses.csv")

head(courses)

# Naming Grade_Distribution
grade_distribution <- grades |> 
  count(grade)

head(grade_distribution)
```

## Exercise 1: Changing Order
```{r}
grade_distribution |> 
  ggplot(aes(x = grade, y = n)) +
    geom_col()
```
The order of the grades is goofy! Construct a new column plot, manually reordering the grades from high (A) to low (NC) with “S” and “AU” at the end:
```{r}
 grade_distribution |>
  mutate(grade = fct_relevel(grade, c("A", "A-", "B+", "B", "B-", "C+", "C", "C-", "D+", "D", "D-", "NC", "S", "AU"))) |>
  ggplot(aes(x = grade, y = n)) +
    geom_col()
```
Construct a new column plot, reordering the grades in ascending frequency (i.e. how often the grades were assigned):
```{r}
grade_distribution |>
  mutate(grade = fct_relevel(grade, c("D-", "D+", "D", "NC", "AU", "C-", "C", "C+", "B-", "S", "B", "B+", "A-", "A"))) |>
  ggplot(aes(x = grade, y = n)) +
    geom_col()
```
Construct a new column plot, reordering the grades in descending frequency (i.e. how often the grades were assigned):
```{r}
grade_distribution |>
  mutate(grade = fct_reorder(grade, n, .desc = TRUE)) |>
  ggplot(aes(x = grade, y = n)) +
    geom_col()
```
## Exercise 2: Changing Factor Level Labels

It may not be clear what “AU” and “S” stand for. Construct a new column plot that renames these levels “Audit” and “Satisfactory”, while keeping the other grade labels the same and in a meaningful order:
```{r}
grade_distribution |>
  mutate(grade = fct_relevel(grade, c("A", "A-", "B+", "B", "B-", "C+", "C", "C-", "D+", "D", "D-", "NC", "S", "AU"))) |>
  mutate(grade = fct_recode(grade, "Satisfactory" = "S", "Audit" = "AU")) |>  
  ggplot(aes(x = grade, y = n)) +
    geom_col()
```



