---
title: "ica-wrangling"
format: html
---

```{r}
# Load tidyverse & data
library(tidyverse)
elections <- read.csv("https://mac-stat.github.io/data/election_2020_county.csv")
```

### Exercise 1: select Practice
```{r}
# Define elections_small
elections_small <- elections |>
select(state_name, county_name, total_votes_20, repub_pct_20, dem_pct_20, total_votes_20, total_votes_16,dem_pct_16)

# Check out the first 6 rows to confirm your code did what you think it did!
head(elections_small)
```
### Exercise 2: Filter Demo
```{r}
# Keep only data on counties in Hawaii
elections_small |>
filter(state_name == "Hawaii")
```

```{r}
# What does this do?
elections_small |> 
filter(state_name %in% c("Hawaii", "Delaware"))
```

This code filters for all counties in Hawaii and Delaware.

```{r}
# Keep only data on counties where the Republican got MORE THAN 93.97% of the vote in 2020
# THINK: What variable is relevant here?

elections_small |> 
filter(repub_pct_20 > 93.97)
```

```{r}
# Keep only data on counties where the Republican got AT LEAST 93.97% of the vote in 2020
# This should have 1 more row (observation) than your answer above

elections_small |> 
filter(repub_pct_20 >= 93.97)
```

```{r}
# Keep only data on counties in Texas where the Democrat got more than 65% of the vote in 2020
# Do this 2 ways.
# Method 1: 2 filters with 1 condition each
elections_small |> 
filter(state_name == "Texas") |> 
filter(dem_pct_20 > 65)

# Method 2: 1 filter with 2 conditions
elections_small |> 
filter(state_name == "Texas", dem_pct_20 > 65)
```

### Exercise 3: Arrange Demo
```{r}
# Arrange the counties in elections_small from lowest to highest percentage of 2020 Republican support
# Print out just the first 6 rows
elections_small |> 
   arrange(repub_pct_20) |> 
   head()
```

```{r}
# Arrange the counties in elections_small from highest to lowest percentage of 2020 Republican support
# Print out just the first 6 rows

elections_small |> 
   arrange(desc(repub_pct_20)) |> 
   head()
```

### Exercise 4: Mutate Demo
## Part a
```{r}
# What did this code do?
elections_small |>
  mutate(diff_20 = repub_pct_20 - dem_pct_20) |>
  head()
```
This code creates a new column with the difference of repub_pct_20 and dem_pct_20.

```{r}
# What did this code do?
elections_small |>
  mutate(repub_votes_20 = round(total_votes_20 * repub_pct_20/100)) |>
  head()
```
This code finds the total votes for Republicans in counties.

```{r}
# What did this code do?
elections_small |>
  mutate(repub_win_20 = repub_pct_20 > dem_pct_20) |>
  head()
```
This code shows the columns where Republicans won over Democrats.

## Part b
```{r}
# You try
# Define a variable that calculates the change in Dem support in 2020 vs 2016
elections_small |>
 mutate(dems_change = dem_pct_20 - dem_pct_16) |>
 head()
```
```{r}
# You try
# Define a variable that determines whether the Dem support was higher in 2020 than in 2016 (TRUE/FALSE)
elections_small |>
 mutate(dem_improve = dem_pct_20 > dem_pct_16) |>
 head()
```

### Exercise 5: Pipe Series
## Part a
```{r}
elections_small |>
  filter(state_name == "Wisconsin",
         repub_pct_20 < dem_pct_20) |>
  arrange(desc(total_votes_20)) |>
  head()
```
## Part b
```{r}
# Now try it. Change the order of filter and arrange below.
elections_small |>
  arrange(desc(total_votes_20)) |>
  filter(state_name == "Wisconsin",
         repub_pct_20 < dem_pct_20) |>
  head()
```
## Part c
So the order of filter() and arrange() did not matter – rearranging them produces the same results. BUT what is one advantage of filtering before arranging?

- An advantage of filtering before arranging is that you can predict what you will produce.

## Part d
```{r}
elections_small |>
  filter(state_name == "Delaware") |>
  mutate(repub_win_20 = repub_pct_20 > dem_pct_20) |>
  select(county_name, repub_pct_20, dem_pct_20, repub_win_20)
```

## Part e
```{r}
# Now try it. Change the order of mutate and select below.
# elections_small |>
#   filter(state_name == "Delaware") |>
#   select(county_name, repub_pct_20, dem_pct_20, repub_win_20)|>
#   mutate(repub_win_20 = repub_pct_20 > dem_pct_20) 
```
There is an error because we are trying to select a variable we don't have.

### Exercise 6: DIY Pipe Series
## Part a
```{r}
elections_small |>
  filter(state_name == "Minnesota") |>
  arrange(desc(dem_pct_20)) |>
  select(county_name,dem_pct_20)
```

## Part b
```{r}
# Define the dataset
# Only store the results once you're confident that they're correct

mn_wi<- elections_small |>
  filter(state_name %in% c("Minnesota", "Wisconsin")) |>
  mutate(dem_change = dem_pct_20 - dem_pct_16) |>
  arrange(dem_change) |>
  select(state_name, county_name,dem_pct_20, dem_pct_16, dem_change)
  
# Check out the first 6 rows to confirm your results
head(mn_wi)
```
## Part c
```{r}
ggplot(mn_wi, aes(x=dem_change,fill = state_name)) +
  geom_density(alpha= .7)
```

### Exercise 7: Summarize Demo 
```{r}
# What does this do?
elections_small |>
  summarize(median(repub_pct_20))
```
This code shows the median percetange of all US counties Republican votes

```{r}
# What does this do?
elections_small |>
  summarize(median_repub = median(repub_pct_20))
```
This code renames the category and shows the median percentage of all US counties Republican vote.

```{r}
# What does this do?
elections_small |>
  summarize(median_repub = median(repub_pct_20), total_votes = sum(total_votes_20))
```
This code adds the category of total votes.

### Exercise 8: summarize + group_by demo
```{r}
elections_small |>
  group_by(state_name)
```

```{r}
# Check out the structure before and after group_by
elections_small |>
  class()

elections_small |>
  group_by(state_name) |>
  class()
```
```{r}
# What does this do?
# (What if we didn't use group_by?)
elections_small |>
  group_by(state_name) |>
  summarize(median_repub = median(repub_pct_20), total_votes = sum(total_votes_20))
```
This code summarizes all 50 states using total votes and median republican votes in counties. If we didn't use group_by, there would be the full 3,109 data points in the data set.
 
### Exercise 9: DIY
## Part a
```{r}
# Sort the *states* from the most to least total votes cast in 2020
elections_small |>
  group_by(state_name) |>
  summarize(total_votes = sum(total_votes_20)) |>
  arrange(desc(total_votes))
```

```{r}
# In 2020, what were the total number of votes for the Democratic candidate and the total number of votes for the Republican candidate in each *state*?

elections_small |>
  group_by(state_name) |>
  summarize(dem_votes = sum(dem_pct_20), repb_votes = sum(repub_pct_20)) 
```
```{r}
# What states did the Democratic candidate win in 2020?
elections_small |>
  group_by(state_name) |>
  mutate(dem_win_20 = repub_pct_20 < dem_pct_20) |>
  select(state_name, repub_pct_20, dem_pct_20, dem_win_20)
```

## Part b
```{r}
# Sort the states from the most to least total votes cast in 2020
# HINT: Calculate the total number of votes in each state, then sort
elections_small |>
  group_by(state_name) |>
  summarize(total_votes = sum(total_votes_20)) |>
  arrange(desc(total_votes))
```

```{r}
# In 2020, what were the total number of votes for the Democratic candidate and the total number of votes for the Republican candidate in each state?
# HINT: First calculate the number of Dem and Repub votes in each *county*
# Then group and summarize these by state
elections_small |>
  group_by(state_name) |>
  summarize(dem_votes = sum(dem_pct_20), repb_votes = sum(repub_pct_20)) 
```

```{r}
# What states did the Democratic candidate win in 2020?
# HINT: Start with the results from the previous chunk, and then keep only some rows

elections_small |>
  group_by(state_name) |>
  mutate(dem_win_20 = repub_pct_20 < dem_pct_20) |>
  select(state_name, repub_pct_20, dem_pct_20, dem_win_20)
```

### Exercise 10: Practice on New Data

```{r}
world_cup <- read.csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-11-29/worldcups.csv")
```

```{r}
# In what years did Brazil win the World Cup?
world_cup |>
  filter(winner == "Brazil") |>
  select(year, winner)
```
```{r}
# What were the 6 World Cups with the highest attendance?
world_cup |>
  arrange(desc(attendance)) |>
  select(year,attendance) |>
  filter(attendance >= 2724604)
```
```{r}
# Construct a univariate plot of goals_scored (no wrangling necessary)
# This provides a visual summary of how the number of goals_scored varies from World Cup to World Cup

ggplot(world_cup, aes(x=goals_scored)) +
  geom_histogram(binwidth = 8)
```

```{r}
# Let's follow up the plot with some more precise numerical summaries
# Calculate the min, median, and max number of goals_scored across all World Cups
# NOTE: Visually compare these numerical summaries to what you observed in the plot

world_cup |>
  summarize(median_goals = median(goals_scored), max_goals = max(goals_scored), min_goals = min(goals_scored))
```
```{r}
# Construct a bivariate plot of how the number of goals_scored in the World Cup has changed over the years
# No wrangling necessary
```

```{r}
# Our above summaries might be a bit misleading.
# The number of games played at the World Cup varies.
# Construct a bivariate plot of how the typical number of goals per game has changed over the years
```

