---
title: "Mulivariate Viz"
---

Use this file for practice with the **mulivariate viz** in-class activity.  Refer to the class website for details.

# 5.1 Notes
```{r}
library(tidyverse)

# Import data
weather <- read.csv("https://mac-stat.github.io/data/weather_3_locations.csv") |> 
  mutate(date = as.Date(date))  

# Check out the first 6 rows
head(weather)
```
# What are the units of observation?
date, location, mintemp, maxtemp,rainfall, evaporation, sunshine, windgustdir, windgustspeed, winddir9am, winddir3pm, windspeed9am, windspeed3pm, humidity9am, humidity3pm, pressure9am, pressure3pm, cloud9am, cloud3pm, temp9am, temp3pm, raintoday, risk_mm,raintomorrow

# How many data points do we have? 


# What type of variables do we have?



# In-Class Activities - Exercises
```{r}
# Import and check out data
education <- read.csv("https://mac-stat.github.io/data/sat.csv")
head(education)

#library
library(tidyverse)
```

## Exercise 1: SAT scores
### Part a
```{r}
ggplot(education, aes(x=sat)) + geom_density() 
```
### Part b
The sat scores are slightly left-skewed and bi-model. I don't know why. SAT scores range from 840 to 1110. Typical outcomes are 900 and 1040. 

## Exercise 2: SAT Scores vs Per Pupil Spending & SAT Scores vs Salaries
### Part a
```{r}
# Construct a plot of sat vs expend
# Include a "best fit linear regression model" (HINT: method = "lm")
ggplot(education, aes(x = expend, y = sat)) + geom_point() + geom_smooth(method = lm)

```
```{r}
# Construct a plot of sat vs salary
# Include a "best fit linear regression model" (HINT: method = "lm")
ggplot(education, aes(x = salary, y = sat)) + geom_point() + geom_smooth(method = lm)
```

### Part b
What are the relationship trends between SAT scores and spending? Is there anything that surprises you?

- I'm surprised that the average SAT scores goes down with salary and expenditure. That doesn't really make sense to me.

## Exercise 3: SAT Scores vs Per Pupil Spending and Teacher Salaries
```{r}
ggplot(education, aes(x = expend, y = sat)) + geom_point() + facet_wrap (~sat)
```
```{r}
ggplot(education, aes(x = salary, y = sat, color = expend)) + geom_point() 
```
```{r}
ggplot(education, aes(x = expend, y = sat, color = salary)) + geom_point() 
```

## Exercise 4: Another way to incoporate scale
```{r}
ggplot(education, aes(y = sat, x = salary, color = cut(expend, 3))) + 
  geom_point() + 
  geom_smooth(se = FALSE, method = "lm")
```
Describe the trivariate relationship between sat, salary, and expend.

- As SAT scores decrease, salary increases. However as expenditure decreaeses, SAT scores decrease. 

## Exercise 5: Finally an Explanation

### Part a
```{r}
ggplot(education, aes(x = fracCat)) + geom_bar()
```

### Part b 

```{r}
ggplot(education, aes(x = sat, fill = fracCat)) + geom_histogram()
```
```{r}
ggplot(education, aes(x = sat, fill= fracCat)) + geom_density()
```

### Part c
```{r}
ggplot(education, aes(x = sat, y = expend, color = fracCat)) + geom_point() + geom_smooth(method = lm)
```

### Part d
It appears that SAT scores decrease with higher expenditure because the expenditure is actually higher proportionately to the amount of students that actually take the SAT. 
