---
title: "Bivariate Viz"
number-sections: true
execute: 
  warning: false
fig-height: 2.75
fig-width: 4.25
fig-env: 'figure'
fig-pos: 'h'
fig-align: center
code-fold: false
---

# Reading Notes

## Review

```{r}
# Import data
survey <- read.csv("https://ajohns24.github.io/data/112/about_us_2024.csv")
```
### How many students have now filled out the survey?
28 students have filled out the survey

### What type of variables do we have?
The variables in the survey are cafe_mac, minutes_to_campus, fave_temp, and hangout

## EXAMPLE 1: Hangout Preferences {-}

```{r}
# Attach a package needed to use the ggplot functions
library(tidyverse)
```

### Make a ggplot
```{r}
ggplot(survey,aes(x=hangout)) + geom_bar(color = "white") + labs(x="hangout", y = "number of students")
```

## EXAMPLE 2: Temperature Preferences {-}

Students were asked about their ideal outdoor temperature, in degrees Celsius. How did they answer? What was the typical response? What was the range in responses? Were there any outliers? Build and interpret 2 plots that help address these questions.

```{r}
ggplot(survey, aes(x=fave_temp)) + geom_histogram(color = "white",binwidth = 5) + labs(x="Favorite Temp", y = "Number of students")
```

```{r}
ggplot(survey, aes(x = fave_temp)) + geom_bar(color = "blue", fill = "blue") + labs(x = "Favorite Temperature", y = "Count")
```

## EXAMPLE 3 {-}
  a.  3pm temperature (response) vs 9am temperature (predictor)
```{r}
data.frame(temp_3pm = c(24, 26, 20, 15, 15, 15), temp_9am = c(14, 18, 15, 13, 11, 11))
```
  b.  3pm temperature (response) vs location (predictor)
```{r}
weather <- data.frame(temp_3pm = c(24, 26, 20, 15, 15, 0, 40, 60, 57, 44, 51, 75),
                      location = rep(c("A", "B"), each = 6))
weather
```

```{r}
ggplot(weather, aes(x = temp_3pm)) +
      geom_density()
```

c.  `rain_today` (the response) and `location` (the predictor)
```{r}
weather <- data.frame(rain_today = c("no", "no", "no", "no", "yes", "no", "yes", "no", "yes", "yes", "no", "yes"),
                        location = c(rep("A", 7), rep("B", 5)))
    weather
```
How might we modify the below bar plot of location to distinguish between days on which it did or didn’t rain?
```{r}
ggplot(weather, aes(x = location)) +
      geom_bar()
```

# Exercises / HW2

```{r}
# Load data
elections <- read.csv("https://mac-stat.github.io/data/election_2020_county.csv")

#library
library(tidyverse)

# Check it out
head(elections)
```

## Exercise 0: Review

### Part a
```{r}
ggplot(elections, aes(x= winner_20, fill = winner_20)) + 
  geom_bar() +
  scale_fill_manual(values = c("blue", "red")) 
```
Follow Up: 
We know that the Joe Biden won the 2020 Election, so how could we visualize population in the data?

### Part b
```{r}
ggplot(elections, aes(x = repub_pct_20)) + geom_histogram(binwidth = 5, fill="red")
```

Follow up: How can we take this information into account with the other graph?

## Exercise 1: Quantitative vs Quantitative Intuition Check
```{r}
ggplot(elections, aes(x = repub_pct_16, y = repub_pct_20)) + 
  geom_point()
```

## Exercise 2: 2 Quantitiative Variables

```{r}
# Set up the plotting frame
# How does this differ than the frame for our histogram of repub_pct_20 alone?
# Both axis are variable and used to compare data
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16))
```

```{r}
# Add a layer of points for each county
# Take note of the geom!
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16)) +
  geom_point()
```

```{r}
# Change the shape of the points
# What happens if you change the shape to another number?
# Changes the shape of the point, 3 = plus signs
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16)) +
  geom_point(shape = 3)
```

```{r}
# YOU TRY: Modify the code to make the points "orange"
# NOTE: Try to anticipate if "color" or "fill" will be useful here. Then try both.
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16)) +
  geom_point(color="orange")
```

```{r}
# Add a layer that represents each county by the state it's in
# Take note of the geom and the info it needs to run!
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16)) +
  geom_text(aes(label = state_abbr))
```

## Exercise 3: Reflect
Summarize the relationship between the Republican support in 2020 and 2016. Be sure to comment on:

the strength of the relationship (weak/moderate/strong) 
There is a strong relationship between Republican support in 2020 and 2016

the direction of the relationship (positive/negative) 
The direction of the relationship is positive. 

outliers (in what state do counties deviate from the national trend? Any ideas why this might be the case?) 
There are no complete outliers in the data set but some TX states deviate further from the trend.

## Exercise 4: Visualizing trend

```{r}
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16)) +
  geom_point() +
  geom_smooth()
```

### Part a
```{r}
 ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16)) +  
  geom_smooth()
```

### Part b
```{r}
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16)) +
  geom_point() +
  geom_smooth(method = "lm")
```

## Exercise 5: Your Turn

```{r}
# Scatterplot of repub_pct_20 vs median_rent
ggplot(elections, aes(y = repub_pct_20, x = median_rent)) +
  geom_point() +
  geom_smooth(method = "lm")


# Scatterplot of repub_pct_20 vs median_age
ggplot(elections, aes(y = repub_pct_20, x = median_age)) +
  geom_point() +
  geom_smooth(method = "lm")

```

## Exercise 6: A Sad Scatterplot
```{r}
ggplot(elections, aes(y = repub_pct_20, x = historical)) +
  geom_point()
```

A scatterplot might not be an effective vizualization because the categorical variable doesn't align with the strengths of a scatterplot.

## Exercise 7: Quantitative vs Categorical – Violins & Boxes

```{r}
# Side-by-side violin plots
ggplot(elections, aes(y = repub_pct_20, x = historical)) +
  geom_violin()
```

```{r}
# Side-by-side boxplots (defined below)
ggplot(elections, aes(y = repub_pct_20, x = historical)) +
  geom_boxplot()
```

*Reflect:*
Republican support is highest in historically red counties, where as Republican support is lower in historically blue counties. However, in 2020 many counties showed less support of Republicans because they are lower outliers

## Exercise 8: Quantitative vs Categorical – Intuition Check

```{r}
ggplot(elections, aes(x = repub_pct_20, fill = historical)) +
  geom_density()
```

*Fixing colors*
```{r}
ggplot(elections, aes(x = repub_pct_20, fill = historical)) +
  geom_density() + 
  scale_fill_manual(values = c("blue", "purple", "red"))
```

## Exercise 9: Quantitative vs Categorical – Density Plots

```{r}
# Colors are wrong and axis labels are bad
ggplot(elections, aes(x = repub_pct_20, fill = historical)) +
  geom_density()
```

```{r}
# Allows you to manually change the colors of the different categories. 
ggplot(elections, aes(x = repub_pct_20, fill = historical)) +
  geom_density() +
  scale_fill_manual(values = c("blue", "purple", "red"))
```

```{r}
# Alpha = 0.5 changes the opacity of the different values on the graph
# Play around with different values of alpha, between 0 and 1
ggplot(elections, aes(x = repub_pct_20, fill = historical)) +
  geom_density(alpha = 0.7) +
  scale_fill_manual(values = c("blue", "purple", "red"))
```

```{r}
# facet_wrap separates the values into seperate graphs. 
ggplot(elections, aes(x = repub_pct_20, fill = historical)) +
  geom_density() +
  scale_fill_manual(values = c("blue", "purple", "red")) +
  facet_wrap(~ historical)
```

```{r}
# Let's try a similar grouping strategy with a histogram instead of density plot.
# This is terrible because it is hard to read and doesn't clearly visualize the data.
ggplot(elections, aes(x = repub_pct_20, fill = historical)) +
  geom_histogram(color = "white") +
  scale_fill_manual(values = c("blue", "purple", "red"))
```

## Exercise 10

*Which do you like best?*
I like the density plots the best because they are easy to read and are very customizable for presenting quantitative data.
*What is one pro of density plots relative to boxplots?*
A pro of density plots relative to box plots is that you can show data compared to one another, through layering rather than just a side-by-side comparison.
*What is one con of density plots relative to boxplots?*
A con of density plots is that they are not as usful for qualatative data.

## Exercise 11: Categorical vs Categorical – Intuition Check

```{r}
# Plot 1: adjust this to recreate the top plot
ggplot(elections, aes(x = historical,fill = winner_20)) +
  geom_bar() 
```

```{r}
# Plot 2: adjust this to recreate the bottom plot
ggplot(elections, aes(x = winner_20)) +
  geom_bar()
```

## Exercise 12: Categorical vs Categorical

```{r}
# A stacked bar plot
# "historical" is the fill color, meaning it shows the visualizes the data by color. "winner_20" is the x-axis so it shows the data represented through the bar plot.
ggplot(elections, aes(x = historical, fill = winner_20)) +
  geom_bar()
```

```{r}
# A faceted bar plot
ggplot(elections, aes(x = winner_20)) +
  geom_bar() +
  facet_wrap(~ historical)
```

```{r}
# A side-by-side bar plot
# Note the new argument to geom_bar
ggplot(elections, aes(x = historical, fill = winner_20)) +
  geom_bar(position = "dodge")
```

```{r}
# A proportional bar plot
# Note the new argument to geom_bar
ggplot(elections, aes(x = historical, fill = winner_20)) +
  geom_bar(position = "fill")
```

### Part a
One pro of using the proportional bar plot is that you can compare the amount of each data point to a scale of 1, making it easier to show similarities or difference. One con of using the proportional bar plot is that it removes the understanding of scale and volume of total count.
### Part b
The side-by-side bar plot is my favorite because it shows volume of count but also visualizes the data in a way that is easy to read and understand partisan lines.

## Exercise 13: Practice
```{r}
weather <- read.csv("https://mac-stat.github.io/data/weather_3_locations.csv")
```

